% =================================================================
%     Acknowledgements
% =================================================================
\selectlanguage{english}
\chapter*{Acknowledgements}



This thesis, completed within a year of very intense experiences, actually spans over the last three or four years of my life. Those were truly crucial years in setting the motivation for coming to Nova and persuing this line of work, now giving its first results. For this, I thank Isa. During all these years she always supported me and believed me, and without her company and encouragement this work wouldn't have happened. She was always available to listen to my wandering thoughts on logic and absurdity, as well as my concerns and fears regarding my work and its success. The moments I spent with her were always beautiful and full of joy and tranquility. Her love and friendship were the best strenghts I could have had to give this first step, and those are things that I \emph{always} keep with me. Thanks.

My mother and my sister are among the most patient persons in the world. They always let me do things my way, even when it interfered with their lives and schedules, and they always beared with my sudden special needs, even when this meant more work for them. My mother was specially inspiring in the way she has always fought everything without ever giving up, a beautiful persistence I'm sure I got from her.

On the academic side, Professor Moniz Pereira made this whole adventure possible. He was the kind of supervisor I wanted, always keeping me occupied with challenges and always available for my doubts and difficulties. The fact that he accepted me as his student and that he believed in my work was of the utmost importance to me. His constant dedication to scientific research is truly inspiring and his contagious sense of humor was essential for the completion of this thesis.

Thanks are also due to the Departamento de Informática and the Centro de Inteligência Artificial, for providing me with working conditions and great examples among the people working there. The Secretariat staff has always been helpful and patient in all the things I required, for which I also thank them. A special word of thanks also goes to the Multimedia and Databases Laboratories research group, at Escola Superior de Tecnologia, Instituto Politécnico de Setúbal, for providing me alternative working conditions.

I also have to thank my juri, Professor Carlos Damásio and Professor Carlos Caleiro, whose insights were more than essential for the final version of this work.

Sérgio, Carmen, Miguel, Hugo Silva and João Leite were of great help in proofreading my thesis, for which I thank them.

I have a great thank to give to Sérgio, who was always there to listen to me talking about my work, even though he had his own concerns with pixels, isosurfaces and volume rendering. He was one of the few people who actually understood this work, contributed with new problems and examples, even spent some time solving some of these issues and yet this had nothing to do with what he was working on. We spent some great relaxing moments in math's bar, just talking and enjoying the \emph{views}. Those \emph{views} were indeed crucial to our mental sanity. Thanks for being a great friend and formidable colleague.

Doca de Santo at Alcântara and Bar Proíbido at Belém were alternative working places. Besides having good working environments, these were also relaxing places to escape from work.

João Leite was also of great help and inspiration ever since I came to Nova. These ideas weren't among his favourites and so our discussions were a great test to my persistence and my belief in my work, which I think only made it stronger. Besides, he was a great professor, a great research partner and an example both in achievements and dedication to his work.

A special thanks also goes to Sophie, who opened my eyes to a lot of important issues during this period of time. Our moments were filled with brevity but set many important things in motion regarding me and this thesis. We ended up sharing \emph{more} than just music, which had a great impact on me and my work.

Miguel, Mité and Carmen were the greatest friends I could have had. Their companion, support and advices were more than essential to me. Carmen has accompanied me since I came to Nova and has always given me important advices, as well as great moments of fun, along with her big Vimi. Mité was a fun companion in all the singing and guitar playing in her house and at Amadeu's. Not everyone could appreciate the greatness of our voices, which only means they are \emph{that} special. Miguel has probably interfered more with my life than any other person. He was my professor, my supervisor, my work partner and is something like a best friend and an older brother to me. He always had time and the moments I spent in his company were always enjoyable and relaxing. Besides all this, he has always been an inspiration in so many ways. Thank you for always being there.

%\vspace{1cm}
And finally there is Maria, who has (a)mused me when I needed it the most, and gave me so many crazy and passionate moments; an intensity I was missing and which I truly needed. Thanks.